---
title: "Midterm project"
author: "Nathalie Fadel"
date: "10/16/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Notes - dataset observations:
47 weeks in the study, days of week organized alphabetically.
1440 "activity" columns, because there are 1440 min/day. 
These are the activity "epochs", the signals measured per minute

Don't include more than 5 figures. 3 is probably optimal.

#Import and clean dataset
```{r}
exercise_data <-
  exercise_data = read_csv(file = "./data/p8105_mtp_data.csv") %>%
gather(key = minutes, value = activity, activity.1:activity.1440) %>%
  mutate(minutes = str_replace(minutes, "activity.", ""),
         minutes = fct_relevel(minutes, str_c("", c("1-1440")))) %>%
  mutate(minutes = as.numeric(minutes)) %>%
  rename(epochs = activity) %>%
  mutate(day = factor(day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))) %>%
  arrange(week, day, minutes)
  
exercise_data = exercise_data %>%
  group_by(week, day) %>%
  mutate(total_exercise = sum(epochs))

exercise_data %>% group_by(week) %>% summarize(mean(total_exercise)) #mean amount of total daily activity for each week

exercise_data %>% 
  group_by(week) %>% 
  summarize(avg_total_exercise = mean(total_exercise)) %>%
  ggplot(aes(x = week, y = avg_total_exercise)) +
  geom_line() +
  labs(y = "average amount of activity", x = "week number")
#scale of y axis is CRAZY, need to scale back by a factor of like, 10^3.

```
This dataset has `r nrow(exercise_data)` observations and `r ncol(exercise_data)` variables. The "epochs" variable is the numer of electric signals recorded per minute of the day, and the "minutes" variable is the minute per day that the epochs were recorded. 

Discuss any additional exploratory analyses you conduct, but only include results you think are interesting (e.g. visually inspect distributions for outliers, but include only if these are informative).
```{r}

```